---
title: "Investigating your shroom for the presence of repeats"
author: "Nathan Brouwer modified by Kayla Neal"
date: "10/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Add the necessary code to make this script function.

## Preliminaries

### Load packages

```{r}
library(seqinr)
library(rentrez)
library(compbio4all)
library(Biostrings)
```



### Get data shroom data

Download sequence of hShroom 3 protein: NP_001166171.1

```{r}
shroom3 <- rentrez::entrez_fetch(id = "NP_001166171.1",
                                      db = "protein", 
                                      rettype="fasta")
```

Clean and set up sequence as vector.

```{r}
shroom3_vector <- fasta_cleaner(shroom3)
```

Confirm set up using str()
```{r}
str(shroom3_vector)
```


Check length - Shroom is on the big end of proteins 
```{r}
length(shroom3_vector)
```

I am therefore going to subset it initially so the example plots don't take as long to make
```{r}
shroom3_vector <- shroom3_vector[1:400]
```



Make a 2 x 2 grid of dotplots to explore effect of changing window size and nmatch
```{r}
# set up 2 x 2 grid, make margins thinner
par(mfrow = c(2,2), 
    mar = c(0,0,2,1))

# plot 1: Shroom - Defaults
dotPlot(shroom3_vector, 
        shroom3_vector, 
        wsize = 1, 
        nmatch = 1, 
        main = "Shroom Defaults")

# plot 2 Shroom - size = 10, nmatch = 1
dotPlot(shroom3_vector, shroom3_vector, 
        wsize = 10, 
        nmatch = 1, 
        main = "Shroom - size = 10, nmatch = 1")

# plot 3: Shroom - size = 10, nmatch = 5
dotPlot(shroom3_vector, shroom3_vector, 
        wsize = 10, 
        nmatch = 5, 
        main = "Shroom - size = 10, nmatch = 5")

# plot 4: size = 20, nmatch = 5
dotPlot(shroom3_vector, shroom3_vector, 
        wsize = 20,
        nmatch =5,
        main = "Shroom - size = 20, nmatch = 5")

# reset par() - run this or other plots will be small!
par(mfrow = c(1,1), 
    mar = c(4,4,4,4))
```


Make ANOTHER  2 x 2 grid of dotplots to explore effect of changing window size and nmatch
```{r}
# set up 2 x 2 grid, make margins thinner
par(mfrow = c(2,2), 
    mar = c(0,0,2,1))

# plot 1: 
dotPlot(shroom3_vector, shroom3_vector, 
        wsize = 30, 
        nmatch = 5 , 
        main = "Shroom 30 / 5")

# plot 2 Shroom
dotPlot(shroom3_vector, shroom3_vector, 
        wsize = 30, 
        nmatch = 10, 
        main = "Shroom - size = 30, nmatch = 10")

# plot 3: Shroom
dotPlot(shroom3_vector, shroom3_vector, 
        wsize = 5, 
        nmatch = 2, 
        main = "Shroom - size = 5, nmatch = 2")

# plot 4:Shroom
dotPlot(shroom3_vector, shroom3_vector, 
        wsize = 12,
        nmatch = 4,
        main = "Shroom - size = 12, nmatch = 4")

# reset par() - run this or other plots will be small!
par(mfrow = c(1,1), 
    mar = c(4,4,4,4))
```



##  Best plot using normal dotplot

This is the most interesting shroom dotplot based on the changes investigated above

```{r}
# be sure to run par - re-run just in case
par(mfrow = c(1,1), 
    mar = c(4,4,4,4))

dotPlot(shroom3_vector, 
        shroom3_vector,
        wsize = 10, 
        nmatch = 5, 
        main = "Shroom window = 10, match = 5")
```




## Full-length plot

```{r}
dot_plot <- function(seq1, seq2, wsize = 1, wstep = 1, nmatch = 1, col = c("white", 
    "black"), xlab = deparse(substitute(seq1)), ylab = deparse(substitute(seq2)), 
    ...) {
  
  # make sure input works
    if (nchar(seq1[1]) > 1) 
        stop("seq1 should be provided as a vector of single chars")
    if (nchar(seq2[1]) > 1) 
        stop("seq2 should be provided as a vector of single chars")
    if (wsize < 1) 
        stop("non allowed value for wsize")
    if (wstep < 1) 
        stop("non allowed value for wstep")
    if (nmatch < 1) 
        stop("non allowed value for nmatch")
    if (nmatch > wsize) 
        stop("nmatch > wsize is not allowed")
  
  # internal function
    mkwin <- function(seq, wsize, wstep) {
        sapply(seq(from = 1, to = length(seq) - wsize + 1, by = wstep), 
            function(i) c2s(seq[i:(i + wsize - 1)]))
    }
    wseq1 <- mkwin(seq1, wsize, wstep)
    wseq2 <- mkwin(seq2, wsize, wstep)
    if (nmatch == wsize) {
        xy <- outer(wseq1, wseq2, "==")
    }
    else {
        "%==%" <- function(x, y) colSums(sapply(x, s2c) == sapply(y, 
            s2c)) >= nmatch
        xy <- outer(wseq1, wseq2, "%==%")
    }
    
    # compile output in list
    out <- list(x = seq(from = 1, to = length(seq1), length = length(wseq1)),
                y = seq(from = 1, to = length(seq2), length = length(wseq2)),
                z = xy)
}

```

Use the new dotplot function defined above on the full-length protein, save output (doesn't autoplot)
```{r}
my_dot_out <- dot_plot(shroom3_vector,
                       shroom3_vector, 
        wsize = 10, 
        wstep = 5,
        nmatch = 5)
```

Get rid of upper triangular portion
```{r}
my_dot_out$z[lower.tri(my_dot_out$z)] <- FALSE

```

Do some weird prep (don't worry about it)
```{r}
my_dot_out$z <- my_dot_out$z[, nrow(my_dot_out$z):1]
```

Plot using image() command. 
```{r}
# seriously - it will drive you crazy if you forget about this
par(mfrow = c(1,1), 
    mar = c(4,4,4,4))

# plot with image()
image(x = my_dot_out$x, 
      y = my_dot_out$y, 
      z = my_dot_out$z, 
      main ="Shroom3" )
```

## Focal subplot

Use new function on the full-length protein, save output (doesn't autoplot).
Subset from 200 to 400
```{r}
my_dot_out <- dot_plot(shroom3_vector[200:400],
                       shroom3_vector[200:400], 
        wsize = 15, 
        wstep = 1,
        nmatch = 5)
```

Get rid of upper triangular portion
```{r}
my_dot_out$z[lower.tri(my_dot_out$z)] <- FALSE

```

Do some weird prep (don't worry about it)
```{r}
my_dot_out$z <- my_dot_out$z[, nrow(my_dot_out$z):1]
```

Plot using image() command.
```{r}
# seriously - it will drive you crazy if you forget about this
par(mfrow = c(1,1), 
    mar = c(4,4,4,4))

# plot with image()
image(x = my_dot_out$x, 
      y = my_dot_out$y, 
      z = my_dot_out$z, 
      main = "Shroom Repeats DotPlot")
```



